Model

### 2019-04-26 stats/model.Rmd
update kids novelty model
hierarchical version of item parameters; begin to sketch out nesting within items and participants

@@

    var sem_slope = uniformDrift({a: -2, b: 2, width: 0.2})
    var sem_int = uniformDrift({a: 0, b: 5, width: 0.2})

@@

      var sem_knowledge = logistic(sem_int + sem_slope * age)

@@

      var modelPredictions = pragmaticListener({label: "novel_word"}, priorProbs, speakerOptimality,sem_knowledge)
      // display(modelPredictions.score(row.correct))
      observe(modelPredictions, row.correct)

@@

    query.add(cndtn, [sem_int, sem_slope])

@@

  //query.add("speaker_optimality", [speakerOptimality])
  query.add("speaker_optimality", [so_int, so_slope])
  return query 
}
'
```

```{r}
me_model_data <- read_csv("../data/child_basic_me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(age = age_num - min(age_num))

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH", 
                        samples = 10000, 
                        burn = 2500,
                        verbose = T,
                        verboseLag = 500)
)

#saveRDS(sem_know_model, "../saves/sem_know_model.rds")
sem_know_model <- readRDS("../saves/sem_know_model.rds")
```

@@

```{r}
ggplot(sem_know, aes(value))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(parameter~item, scales = 'free')+
  theme_few()
```

@@

inv_model_plot <- ggplot(inv_predictions, aes(x=age,y = me_effect, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Inverse Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()
```

@@

data_plot <- ggplot(data = p2, aes(x = age_num, y = correct, col = familiar, fill = familiar)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  geom_smooth(method = "loess", se = F, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(~familiar)+
  ggtitle("Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")+
  scale_fill_viridis_d(name = "familiar object")

@@

>>>>

  var sem_slope_global = uniformDrift({a: -2, b: 2, width: 0.2})
  var sem_int_global = uniformDrift({a: 0, b: 5, width: 0.2})

  var item_slope_sigma = uniformDrift({a: 0, b: 5, width: 0.2})
  // conceptually it might not make sense to have two sigmas, but maybe it does
  // intercept sigma probably will be smaller than slope sigma
  var item_int_sigma = uniformDrift({a: 0, b: 5, width: 0.2})

@@

    var sem_slope_item = gaussianDrift({mu: sem_slope_global, sigma: item_slope_sigma, width: 0.5})
    var sem_int_item = gaussianDrift({mu: sem_int_global, sigma: item_int_sigma, width: 0.5})

    // var sem_slope = uniformDrift({a: -2, b: 2, width: 0.2})
    // var sem_int = uniformDrift({a: 0, b: 5, width: 0.2})

@@

      var sem_knowledge = logistic(sem_int_item + sem_slope_item * age)

@@

      var modelPredictions = pragmaticListener({label: "novel_word"}, 
            priorProbs, speakerOptimality, sem_knowledge)
      // display(modelPredictions.score(row.correct))

      observe(modelPredictions, row.correct)

      query.add(["modelPrediction", cndtn], [age, Math.exp(modelPredictions.score(1))])

@@

    query.add(["parameter", cndtn], [sem_int_item, sem_slope_item])

@@

  query.add(["parameter", "speaker_optimality"], [so_int, so_slope])
  query.add(["parameter", "global_sem"], [sem_int_global, sem_slope_global])
  query.add(["parameter", "global_sem_sigmas"], [item_int_sigma, item_slope_sigma])

  return query 
}
'
```

```{r}
# me_model_data <- read_csv("../data/child_basic_me.csv")%>%
#     filter(trial != "train1",
#            trial != "train2")%>%
#   mutate(age = age_num - min(age_num))

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 1,
  cores = 1,
  inference_opts = list(method = "incrementalMH", 
                        samples = 10, 
                        burn = 5,
                        verbose = T)#,
                        #verboseLag = 500)
)

#saveRDS(sem_know_model, "../saves/sem_know_model.rds")
sem_know_model <- readRDS("../saves/sem_know_model.rds") %>%
  filter(Chain != 1)
```

@@

```{r}
ggplot(sem_know, aes(x = value))+#, fill = factor(Chain)))+
  geom_density(alpha = 0.3)+
  xlab("Parameter")+
  facet_grid(parameter~item, scales = 'free')+
  theme_few()
```

@@

inv_model_plot <- ggplot(inv_predictions, aes(x=age,y = me_effect, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Inverse Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()

inv_model_plot
```

@@

data_plot <- ggplot(data = p2, aes(x = age_num, y = correct, col = familiar, fill = familiar)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  geom_smooth(method = "loess", se = F, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(~familiar)+
  ggtitle("Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")+
  scale_fill_viridis_d(name = "familiar object")
data_plot
```

@@

#### WORK IN PROGRESS: nesting within participant

to do:
- bookkeeping: index correct participant/item variables at relevant part of code
- inference: try HMC, variational

```{r work in progress}
semStr <- '
var data = dataFromR.data

  // global parameters
  var so_slope = uniformDrift({a: -2, b: 2, width: 0.2})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.2})

  var sem_slope_global = uniformDrift({a: -2, b: 2, width: 0.2})
  var sem_int_global = uniformDrift({a: 0, b: 5, width: 0.2})

  var item_slope_sigma = uniformDrift({a: 0, b: 5, width: 0.2})
  var item_int_sigma = uniformDrift({a: 0, b: 5, width: 0.2})

  var participant_age_adjust = gaussianDrift({mu: 0, sigma: 1, width: 0.2})

  var sem_slope_item = repeat(familiars.length, function(){
    gaussianDrift({mu: sem_slope_global, sigma: item_slope_sigma, width: 0.5} )
  })

  var sem_int_item = gaussianDrift({mu: sem_int_global, 
        sigma: item_int_sigma, width: 0.5})

      var sem_knowledge = logistic(sem_int_item + sem_slope_item * age)

      var modelPredictions = pragmaticListener({label: "novel_word"}, 
            priorProbs, speakerOptimality, sem_knowledge)
      // display(modelPredictions.score(row.correct))

      observe(modelPredictions, row.correct)

      query.add(["modelPrediction", cndtn], [age, Math.exp(modelPredictions.score(1))])

    query.add(["parameter", cndtn], [sem_int_item, sem_slope_item])

  query.add(["parameter", "speaker_optimality"], [so_int, so_slope])
  query.add(["parameter", "global_sem"], [sem_int_global, sem_slope_global])
  query.add(["parameter", "global_sem_sigmas"], [item_int_sigma, item_slope_sigma])

  return query 
}
'
```
